---
title: "p8105_hw6_ms5965"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(ggplot2)
library(readxl)
library(modelr)
library(mgcv)


knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

```{r}
homicide_df = 
  read_csv("data/homicide_data.csv", na = c("", "NA", "Unknown")) %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) %>% 
  filter(
    victim_race %in% c("White", "Black"),
    city_state != "Tulsa, AL") %>% 
  select(city_state, resolution, victim_age, victim_race, victim_sex)
```

Baltimore, MD

```{r}
baltimore_df =
  homicide_df %>% 
  filter(city_state == "Baltimore, MD")
glm(resolution ~ victim_age + victim_race + victim_sex, 
    data = baltimore_df,
    family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(term, OR, starts_with("CI")) %>% 
  knitr::kable(digits = 3)
```

All cities:

```{r}
models_results_df = 
  homicide_df %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = 
      map(.x = data, ~glm(resolution ~ victim_age + victim_race + victim_sex, data = .x, family = binomial())),
    results = map(models, broom::tidy)
  ) %>% 
  select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(city_state, term, OR, starts_with("CI")) 
```

```{r}
models_results_df %>% 
  filter(term == "victim_sexMale") %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


## Problem 2

```{r cleaning df}
bw_df = 
  read_csv("data/birthweight.csv", na = c("", "NA", "Unknown")) %>% 
  janitor::clean_names() %>% 
  mutate(
    babysex = as.factor(babysex), 
    frace = as.factor(frace),
    mrace = as.factor(mrace),
    malform = as.factor(malform),
    parity = as.factor(parity),
    pnumlbw = as.factor(pnumlbw),
    pnumsga = as.factor(pnumsga)
  )
```

```{r}
bw_model_1 = lm(bwt ~ gaweeks + smoken, data = bw_df, na.action = na.exclude) 
```

```{r plots}
bw_df %>% 
  modelr::add_residuals(bw_model_1) %>%
  modelr::add_predictions(bw_model_1) %>% 
  ggplot(aes(x = gaweeks, y = resid)) + 
  geom_violin() +
  labs(
    x = "Gestational age in weeks",
    y = "Residuals"
  )
```

Baby's birth weight was modeled against gestational age in weeks and mother's smoking habits in a linear regression model. Smoking can act as a confounder given that it can result in preterm births, hence, lower gestational age. It can also impact birth outcomes i.e. birth weight. Residuals were analyzed to identify issues in model fit. 


Models 2 & 3:

```{r}
bw_model_2 = lm(bwt ~ gaweeks + blength, data = bw_df, na.action = na.exclude) 

bw_model_3 = lm(bwt ~ bhead*babysex*blength, data = bw_df, na.action = na.exclude) 
```

```{r}
bw_df %>% 
  gather_predictions(bw_model_1, bw_model_2, bw_model_3) %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = gaweeks, y = bwt)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red") + 
  facet_wrap(~model)

```

```{r}
set.seed(1)

cv_df = 
  crossv_mc(bw_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)) %>% 
  mutate(
    bw_model_1 = map(train, ~ lm(bwt ~ gaweeks + smoken, na.action = na.exclude, data = .x)),
    bw_model_2 = map(train, ~ lm(bwt ~ gaweeks + blength, na.action = na.exclude, data = .x)),
    bw_model_3 = map(train, ~ lm(bwt ~ bhead*babysex*blength, na.action = na.exclude, data = .x))
    ) %>%   
  mutate(
    rmse_model1 = map2_dbl(bw_model_1, test, ~rmse(model = .x, data = .y)),
    rmse_model2 = map2_dbl(bw_model_2, test, ~rmse(model = .x, data = .y)),
    rmse_model3 = map2_dbl(bw_model_3, test, ~rmse(model = .x, data = .y))) %>% 
    select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
  
cv_df
```

